Active Learning for SAT Solver Benchmarking

### 2022-10-10 paper/paper.tex
Rewrite section 5.1

\section{Experimental Design}

@@

\section{Evaluation}
\label{sec:eval}
Now that we are all set, we want to, first, discuss preliminary results including the choice of the runtime-discretization method and the choice of the machine-learning predictor.
Thereafter, we look at end-to-end results and instance-family importance.
%JB: instance families not really motivated till now, only appear in abstract and in text regarding stratification

\subsection{Preliminary Results}
\label{sec:evalprel}
%JB: if we dedicate a whole section to these results, we should probably rename this section (not "preliminary" in the typical sense); for TACAS, I guess this can be cut a bit
An exhaustive grid search of all approaches outlined in Section~\ref{sec:main} is infeasible.
We are, therefore, narrowing down the search by filtering based on preliminary results.
Since all relevant decisions, i.e., \emph{selection}, \emph{stopping}, and \emph{ranking}, revolve around the choice of prediction model, we are first looking at the \emph{Matthews Correlation Coefficient} (MCC) scores~\cite{gorodkin2004comparing,matthews1975comparison} that a particular prediction model achieves.
MCC scores are great in dealing with class imbalances in contrast to conventional metrics like accuracy.

\begin{table}
  \centering
  \caption{
  	Time-out prediction Matthews Correlation Coefficient performances.
  	Results are obtained using cross-validation.
  	We provide \SI{64}{\%} of ground-truth for training, use \SI{16}{\%} as a validation set, and \SI{20}{\%} as a test set (5-fold cross-validation).
  }
  \label{tab:mcc}
  \begin{tabular}{lc}
    Prediction Model & Avg. MCC (Std.) \\
    \hline
    Stacking ensemble (QDA, RF)           & 0.9527 (0.0292) \\
    Quadratic-discriminant analysis (QDA) & 0.9290 (0.0339) \\
    Random forest (RF)                    & 0.8530 (0.0479) \\
    AdaBoost with decision trees          & 0.8384 (0.0444) \\
    AdaBoost with logistic regression     & 0.8158 (0.0985) \\
    Decision tree                         & 0.8059 (0.0707) \\
    Logistic regression                   & 0.8052 (0.1018) \\
    kNN classifier                        & 0.7885 (0.1521) \\
    MLP classifier                        & 0.7760 (0.1408) \\
    Support-vector classifier             & 0.7757 (0.2149) \\
    Naive Bayes                           & 0.7306 (0.1394) \\
    \hline
  \end{tabular}
\end{table}

@@

To obtain the results in Table~\ref{tab:mcc}, we perform cross-validation among all solvers: we assume that all other solver runtimes are known and provide \SI{64}{\%} of ground-truth runtime labels of the target solver for training, use \SI{16}{\%} as a validation set for hyper-parameter tuning, and \SI{20}{\%} as a test set (5-fold cross-validation).
This is repeated for each solver once.
Note that we perform cross-validation on two levels, i.e., among the solvers and the target solver runtimes.
We report the average performances on the test sets.
Since the differences in the MCC-performance distributions of the best and second-best approaches are statistically significant regarding a Wilcoxon signed-rank test with $\alpha = \SI{5}{\%}$, we only use the best approach for all further experiments as a prediction model.

As a runtime discretization technique, we use hierarchical clustering with $k = 3$ using a log-single-link criterion for all further experiments.
Each non-time-out runtime starts in a separate interval.
We then gradually merge intervals whose single-link logarithmic distance is the smallest until the desired number of partitions is reached.
As already mentioned in Section~\ref{sec:main-model}, this approach has good properties regarding the predictiveness of the PAR-2 score ranking and its discriminatory power.
Also, preliminary experiments showed that our stacking ensemble performed particularly well for this kind of discretization method.
Other clustering approaches that we have tried include hierarchical clustering with mean-, median- and complete-link criteria, as well as $k$-means and spectral clustering.
Predicting runtime labels in a similar experiment as above produced an MCC score of on average \SI{88.82}{\%} ($\sigma = \SI{4.75}{\%}$; in comparison to \SI{95.27}{\%} for predicting time-outs).
%JB: "similar experiment as above" - reference unclear; what do you want to say?

>>>>

\section{Experimental Design}
\label{sec:exdesign}

@@

\section{Evaluation}
\label{sec:eval}
Before evaluating our approach end-to-end as described in Section~\ref{sec:exdesign}, we discuss the choice of the runtime-discretization method and the machine-learning predictor.
Thereafter, we look at the results of our active-learning approach.
Selection decisions of the aforementioned approach also reveal the importance of different problem-instance families to our model.
Instance families comprise instances that are derived from the same application domain, e.g., planning, cryptography, etc., and are a useful tool for both analyzing solver performance and portfolio creation.
%JB: instance families not really motivated till now, only appear in abstract and in text regarding stratification
%TF: Added a sentence for clarification. We could introduce more details in section 4.3.

\subsection{Runtime Prediction}
\label{sec:evalprel}
%JB: if we dedicate a whole section to these results, we should probably rename this section (not "preliminary" in the typical sense); for TACAS, I guess this can be cut a bit
%TF: cut the section down a bit, renamed to "runtime prediction"
An exhaustive grid search of all active-learning approaches outlined in Section~\ref{sec:main} and different runtime-prediction models is infeasible.
Therefore, we use the runtime-prediction model that produced the best results for all further experiments.
We compare runtime predictors by looking at the \emph{Matthews Correlation Coefficient} (MCC) scores~\cite{gorodkin2004comparing,matthews1975comparison} of the classification task.
MCC scores are great in dealing with class imbalances in contrast to conventional metrics like accuracy.

% \begin{table}
%   \centering
%   \caption{
%   	Time-out prediction Matthews Correlation Coefficient performances.
%   	Results are obtained using cross-validation.
%   	We provide \SI{64}{\%} of ground-truth for training, use \SI{16}{\%} as a validation set, and \SI{20}{\%} as a test set (5-fold cross-validation).
%   }
%   \label{tab:mcc}
%   \begin{tabular}{lc}
%     Prediction Model & Avg. MCC (Std.) \\
%     \hline
%     Stacking ensemble (QDA, RF)           & 0.9527 (0.0292) \\
%     Quadratic-discriminant analysis (QDA) & 0.9290 (0.0339) \\
%     Random forest (RF)                    & 0.8530 (0.0479) \\
%     AdaBoost with decision trees          & 0.8384 (0.0444) \\
%     AdaBoost with logistic regression     & 0.8158 (0.0985) \\
%     Decision tree                         & 0.8059 (0.0707) \\
%     Logistic regression                   & 0.8052 (0.1018) \\
%     kNN classifier                        & 0.7885 (0.1521) \\
%     MLP classifier                        & 0.7760 (0.1408) \\
%     Support-vector classifier             & 0.7757 (0.2149) \\
%     Naive Bayes                           & 0.7306 (0.1394) \\
%     \hline
%   \end{tabular}
% \end{table}

@@

%To obtain the results in Table~\ref{tab:mcc},
We perform cross-validation among all solvers: we assume that all other solver runtimes are known and provide \SI{64}{\%} of ground-truth runtime labels of the target solver for training, use \SI{16}{\%} as a validation set for hyper-parameter tuning of the runtime predictor, and \SI{20}{\%} as a test set (5-fold cross-validation).
This is repeated for each solver once.
Note that we perform cross-validation on two levels, i.e., among the solvers and the target solver runtimes.
We report the average performances on the test sets.

ML classification models that we have tried include random forests, decision trees, quadratic-discriminant analysis, AdaBoost, logistic regression, $k$-nearest neighbor, multi-layer perceptrons, support vector machines, and naive Bayes.
A stacking ensemble consisting of quadratic-discriminant analysis and random forests performed best.
It produces an MCC score of \SI{0.9527}{} for time-out prediction and \SI{0.8882}{} for runtime-label prediction.
The results are significantly better than any other model that we have tried regarding a Wilcoxon signed-rank test with $\alpha = \SI{5}{\%}$.

%Since the differences in the MCC-performance distributions of the best and second-best approaches are statistically significant regarding a Wilcoxon signed-rank test with $\alpha = \SI{5}{\%}$, we only use the best approach for all further experiments as a prediction model.

For runtime discretization, we use hierarchical clustering with $k = 3$ using a log-single-link criterion for all further experiments.
Each non-time-out runtime starts in a separate interval.
We then gradually merge intervals whose single-link logarithmic distance is the smallest until the desired number of partitions is reached.
As already mentioned in Section~\ref{sec:main-model}, this approach has good properties regarding the predictiveness of the PAR-2 score ranking and its discriminatory power.
Also, experiments showed that our stacking ensemble performed particularly well for this kind of discretization method.
Other clustering approaches that we have tried include hierarchical clustering with mean-, median- and complete-link criteria, as well as $k$-means and spectral clustering.
%Predicting runtime labels in a similar experiment as above produced an MCC score of on average \SI{88.82}{\%} ($\sigma = \SI{4.75}{\%}$; in comparison to \SI{95.27}{\%} for predicting time-outs).
%JB: "similar experiment as above" - reference unclear; what do you want to say?
%TF: Removed the sentence
